---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 1B - Bodyfat

```{r "setup", include=FALSE}
## set working directory
require("knitr")
library(car)
require("car")
```

The dataset bodyfat contains measurements of bodyfat for a number of men and women 


Variable name | Description
------------- | -------------
gender | gender (male/femal)
fatpct | Measurement of fat percentage


## 1. Can the data in each group (gender) be assumed to be normally distributed

Start by loading the data (make sure datafile is present in working directory)
```{r}
bodyfat <- read.table("bodyfat.txt", header=TRUE)
```

Use both visual aids as well as normality tests to check the normality of fat-percentages on each gender.
```{r}
# Split data according to gender
bodyfat2 <- split(bodyfat$fatpct, bodyfat$gender)
names(bodyfat2) <- c("female","male")

# Check for normality (both samples) with Q-Q Plots
par(mfrow=c(1,2))
qqnorm(bodyfat2$female, main="Q-Q Plot for Females")
qqline(bodyfat2$female)
qqnorm(bodyfat2$male, main="Q-Q Plot for Males")
qqline(bodyfat2$male)

# Plot with confidence intervals
qqPlot(bodyfat2$male)
qqPlot(bodyfat2$female)

# Normality test (Shapiro)
shapiro.test(bodyfat2$male)
shapiro.test(bodyfat2$female)
```

```{r}
# Visual normality test with ecdf and cdf

# Male parameters
xSeqMale <- seq(min(bodyfat2$male), max(bodyfat2$male), by=0.1)
muMale <- mean(bodyfat2$male)
sMale <- sd(bodyfat2$male)

# Female parameters
xSeqFemale <- seq(min(bodyfat2$female), max(bodyfat2$female), by=0.1)
muFemale <- mean(bodyfat2$female)
sFemale <- sd(bodyfat2$female)

# Plot ecdf and cdf
par(mfrow=c(1,2))
plot(ecdf(bodyfat2$male), main="Male")
lines(xSeqMale, pnorm(xSeqMale, muMale, sMale), col="red")
plot(ecdf(bodyfat2$female), main="Female")
lines(xSeqFemale, pnorm(xSeqFemale, muFemale, sFemale), col="red")
```

Both male and female are normally distributed since p-values are > $\alpha$ = 0.05, thus we cannot reject the null hypothesis. 

## 2. Is there a difference in the percentage of body fat for men and women? Perform a t-test
```{r }
t.test(bodyfat2$male, bodyfat2$female)
```

There's no difference in the percentage of body fat for men and women since the t-value is small and the p-value is > $\alpha$ = 0.05, thus we cannot reject the null hypothesis, i.e. the probability that the means of both groups are not statistically different.

## 3. Is there a difference in the percentage of body fat for men and women? Perform a non-parametric test
```{r}
wilcox.test(bodyfat2$female, bodyfat2$male, conf.int = TRUE)
```

There's no difference in the percentage of body fat for men and women since the p-value is > $\alpha$ = 0.05, thus we cannot reject the null hypothesis, i.e. we accept that both groups' means are not statistically different.