---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 1A - Climate


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("ggplot2")
```


The dataset clima contains yearly average temperatures in Denmark and Greenland for the last
hundred years.  

Variable name | Description
------------- | -------------
time | Year
denmark | Average temperature in Denmark
greenland | Average temperature in Greenland

## 1. Plot the two time series and compute the correlation. What is the conclusion?

Start by loading the data (make sure datafile is present in working directory)
```{r}
clima <- read.table("clima.txt", header=TRUE)
```

Use the plot command in R to plot data
```{r}
par(mfrow=c(1,2))
plot(clima$time, clima$denmark, main="Denmark", type="l", xlab="Year", ylab="Temperature (°C)", col="red")
plot(clima$time, clima$greenland, main="Greenland", type="l", xlab="Year", ylab="Temperature (°C)", col="dark green")
```

Plot Temperatures of Greenland and Denmark up against eachother to explore the relation between these two
```{r }
# Line plot
plot(clima$time, clima$denmark, type="l", main="Yearly average temperatures in Denmark and Greenland", col="red", xlab="Year", ylab="Temperature (°C)", ylim=c(-15,10), lwd=2)
lines(clima$time, clima$greenland, col="dark green", lwd=2)
legend(1900,10, legend=c("Denmark", "Greenland"), col=c("red", "dark green"), lty=1:1, cex=1)
correlation=cor(clima$denmark, clima$greenland)
cor.test(clima$denmark, clima$greenland)

# Scatter plot
plot(clima$denmark, clima$greenland, main="Correlation between Denmark and Greenland", xlab="Denmark", ylab="Greenland")

```

## 2 Is the temperature in Denmark increasing when considering the time series 1900-1960? 
```{r}
clima1900 <- clima[clima$time<1961,]
plot(clima1900$time, clima1900$denmark, main="Yearly average temperature in Denmark (1900 - 1960)", type="l", col="red", xlab="Year", ylab="Temperature (°C)", lwd=2)
abline(lm(clima1900$denmark~clima1900$time))
```

No, the temperature is decreasing.

## 3 Is the temperature in Denmark increasing when considering the time series 1960-2000?
```{r}
clima1960 <- clima[clima$time>1959,]
plot(clima1960$time, clima1960$denmark, main="Yearly average temperature in Denmark (1960 - 2000)", type="l", col="dark green", xlab="Year", ylab="Temperature (°C)", lwd=2)
abline(lm(clima1960$denmark~clima1960$time))
```

Yes, the temperature is increasing.