---
output:
  html_document: default
  pdf_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 1A - Climate


```{r "setup", include=FALSE}
## set working directory
require("knitr")
```


The dataset clima contains yearly average temperatures in Denmark and Greenland for the last
hundred years.  

Variable name | Description
------------- | -------------
time | Year
denmark | Average temperature in Denmark
greenland | Average temperature in Greenland

## 1. Plot the two time series and compute the correlation. What is the conclusion?

Start by loading the data (make sure datafile is present in working directory)
```{r}
# Put code here
```

Use the plot command in R to plot data
```{r}
```

Plot Temperatures of Greenland and Denmark up against eachother to explore the relation between these two
```{r }
```

## 2 Is the temperature in Denmark increasing when considering the time series 1900-1960? 
```{r}
```


## 3 Is the temperature in Denmark increasing when considering the time series 1960-2000?

