---
output:
  html_document: default
  pdf_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 1C - calcium


```{r "setup", include=FALSE}
## set working directory
require("knitr")
library(car)
require("car")
```

Does increasing calcium intake reduce blood pressure? Observational studies suggest that there is a
link, and that it is strongest in African-American men. Twenty-one African-American men participated
in an experiment to test this hypothesis. Ten of the men took a calcium supplement for 12 weeks while
the remaining 11 men received a placebo. Researchers measured the blood pressure of each subject
before and after the 12-week period. The experiment was double-blind. The dataset calcium contains
data from the experiment.


Variable name | Description
------------- | -------------
treatment | whether subject received calcium or placebo
begin | aseated systolic blood pressure before treatment
end | aseated systolic blood pressure after treatment
decrease | decrease in blood pressure (begin-end)

## 1. What statistical test is appropriate for comparing the change in blood pressure between the treatment and placebo groups?

Two sample t-test if the samples are normally distributed, if not, we should use Wilcoxon rank sum test.

## 2 .May the data in each group be considered as being normally distributed?

Start by loading data:
```{r}
calcium <- read.table("calcium.txt", header=TRUE)
```

Investigate normality
```{r}
# Split data according to gender
calcium2 <- split(calcium$Decrease, calcium$Treatment)
names(calcium2) <- c("Calcium","Placebo")

# Check for normality (both samples) with Q-Q Plots
par(mfrow=c(1,2))
qqnorm(calcium2$Calcium, main="Q-Q Plot for Calcium")
qqline(calcium2$Calcium)
qqnorm(calcium2$Placebo, main="Q-Q Plot for Placebo")
qqline(calcium2$Placebo)

# Plot with confidence intervals
qqPlot(calcium2$Calcium, ylab="Blood Pressure", main="Calcium")
qqPlot(calcium2$Placebo, ylab="Blood Pressure", main="Placebo")

# Normality test (Shapiro)
shapiro.test(calcium2$Calcium)
shapiro.test(calcium2$Placebo)
```

```{r}
# Visual normality test with ecdf and cdf

# Calcium parameters
xSeqCalcium <- seq(min(calcium2$Calcium), max(calcium2$Calcium), by=0.1)
muCalcium <- mean(calcium2$Calcium)
sCalcium <- sd(calcium2$Calcium)

# Placebo parameters
xSeqPlacebo <- seq(min(calcium2$Placebo), max(calcium2$Placebo), by=0.1)
muPlacebo <- mean(calcium2$Placebo)
sPlacebo <- sd(calcium2$Placebo)

# Plot ecdf and cdf
par(mfrow=c(1,2))
plot(ecdf(calcium2$Calcium), main="Calcium")
lines(xSeqCalcium, pnorm(xSeqCalcium, muCalcium, sCalcium), col="red")
plot(ecdf(calcium2$Placebo), main="Placebo")
lines(xSeqPlacebo, pnorm(xSeqPlacebo, muPlacebo, sPlacebo), col="red")
```

Both calcium and placebo are normally distributed since p-values are > $\alpha$ = 0.05, thus we cannot reject the null hypothesis.

## 3. Test whether the variance in each group can be assumed to be the same


```{r}
varCalcium <- var(calcium2$Calcium)
varPlacebo <- var(calcium2$Placebo)

x<-calcium$End[calcium$Treatment=='Calcium']
y<-calcium$End[calcium$Treatment=='Placebo']
var.test(x,y)
```



## 4. Make a graphical comparison of the treatment means


```{r}
par(mfrow=c(1,2))
boxplot(calcium2$Calcium, main="Calcium", ylim=c(-15,20))
points(mean(calcium2$Calcium), col="red", pch=4, lwd=2)
boxplot(calcium2$Placebo, main="Placebo", ylim=c(-15,20))
points(mean(calcium2$Placebo), col="red", pch=4, lwd=2)
```

## 5. Make the statistical test for comparing the change in blood pressure between the treatment and placebo groups. What is your conclusion? What is the p-value of the test?  


```{r}
t.test(calcium2$Calcium, calcium2$Placebo)
```

Given the data we have, we can say that the mean change in blood pressure between the two groups is not statistically significant, therefore the group's means are similar. This is given by the p-value being > $\alpha$ = 0.05.

## 6. Which non-parametric test could be used if data cannot be assumed to be normally distributed?

We could use Wilcoxon rank sum test.

```{r}
wilcox.test(calcium2$Calcium, calcium2$Placebo, conf.int = TRUE)
```
