---
output:
  html_document: default
  pdf_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 1D - laborforce


```{r "setup", include=FALSE}
## set working directory
require("knitr")
library(car)
require("car")
```

This dataset contains the labor force participation rate (LFPR) of women in 19 cities in the United
States in each of two years (1968 and 1972). The data help to measure the growing presence of women
in the labor force over this period.

Variable name | Description
------------- | -------------
city | City in the US
1972 | Labor Force Participation rate of women in 1972
1968 | Labor Force Participation rate of women in 1968

## 1. Compare LFPR rates in the two years with a pooled t-test since the United States did not change much from 1968 to 1972
Start by loading data:
```{r}
labor <- read.table("labor.txt", header=TRUE)
```

If we suspect that the variance is roughly the same for the groups of 1968 and 1972 we can use a Pooled t-test in R

```{r}
s1968=var(labor$x1968)
s1972=var(labor$x1972)
t.test(labor$x1968, labor$x1972, var.equal=TRUE)
```


## 2. Compare LFPR rates in the two years with a paired t-test

Paired T-test in R
```{r}
t.test(labor$x1968, labor$x1972, paired=TRUE)
```

Difference in the mean labor fource participation of women in both years is statistically significant. Therefore, we reject the null hypothesis and accept the alternative hypothesis where the true difference in means is not equal to 0, i.e. they're different.

## 3. Which of the two methods you have used do you find most appropriate?

Paired t-test gives us a change of the labor force in two different years, it should be the same sample in different times.

## 4. If you have found a significant change - how big is this change then?

It's as big as 2.4577, i.e. the t-crit of the paired t-test.