---
output:
  html_document: default
  pdf_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 4C - Popular


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("dplyr")
```
Subjects were students in grades 4-6 from three school districts in Ingham and Clinton Counties,
Michigan. Chase and Dummer stratified their sample, selecting students from urban, suburban, and
rural school districts with approximately 1/3 of their sample coming from each district. Students
indicated whether good grades, athletic ability, or popularity was most important to them. They also
ranked four factors: grades, sports, looks, and money, in order of their importance for popularity. The
questionnaire also asked for gender, grade level, and other demographic information.

Variable name | Description
------------- | --------------------------------------------
Gender | Boy or girl
Grade | 4, 5, or 6
Age | Age in years
Race | White, Other
Urban/Rural | Rural, Suburban, or Urban school district
School | Brentwood Elementary, Brentwood Middle, Ridge, Sand, Eureka, Brown, Main Portage, Westdale Middle  
Goals |  Student's choice in the personal goals question where options were 1 = Make Good Grades 2 = Be Popular, 3 = Be Good in Sports  
Grades | Rank of "make good grades"(1=most important for popularity, 4=least important)
Sports | Rank of "being good at sports"(1=most important for popularity, 4=least important)
Looks | Rank of "being handsome or pretty"(1=most important for popularity, 4=least important)
Money | Rank of "having lots of money"(1=most important for popularity, 4=least important)

## 1. Analyze the relationship between gender and goals.
```{r}
df <- read.table("popular.txt", header=TRUE, sep="\t")
df1 <- select(df,Gender,Goals)
matrix1 <- table(df1)

# Chi-squared
chi1 <- chisq.test(matrix1)
chi1$expected
chi1

# Visualization
mosaicplot(matrix1, shade=TRUE)

# Fisher test
# fisher.test(matrix)
```

We fail to reject $H_0$ since p-value is less than $\alpha$, i.e. `chi1$p.value` < 0.05. Therefore, variables gender and goals are dependant.

## 2. Analyze the relationship between age and goals.

```{r}
df2 <- select(df,Age,Goals)
matrix2a <- table(df2)

# Chi-squared
chi2 <- chisq.test(matrix2a)
chi2$expected
chi2

# Fisher test
# fisher.test(matrix2)

matrix2b <- matrix2a[-c(2,6),]
matrix2b[1,] <- matrix2a[1,]+matrix2a[2,]
matrix2b[4,] <- matrix2a[5,]+matrix2a[6,]
dimnames(matrix2b)[[1]] <- c("7:9","10","11","12:13")

# Chi-squared
chi2b <- chisq.test(matrix2b)
chi2b$expected
chi2b
mosaicplot(matrix2b, shade=TRUE)
```

They're not dependant.

## 3. Could you suggest other analysis that may be interesting?

```{r}
df3 <- select(df,Urban.Rural,Goals)
matrix3 <- table(df3)

# Chi-squared
chi3 <- chisq.test(matrix3)
chi3$expected
chi3
mosaicplot(matrix3, shade=TRUE)
```

Rural students are under-represented in grades and over-represented in sports.