---
output:
  html_document: default
  pdf_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 4A - Caffeine


```{r "setup", include=FALSE}
## set working directory
require("knitr")
```

In a study the caffeine consumption by marital status among women giving birth was investigated.
The results are given in the table below

Martial-status/    Caffeine | 0  | 1-150 |  151-300 | > 300   
---- | ---- |  ---- | ---- |   ---- | 
Married | 652 | 1537 | 598 | 242
Prev. Married | 36 | 46 | 38 | 21 |
Single | 218 | 327 | 106 | 67 |

## 1. Type the data into R so that appropriate analysis can be carried out
```{r}
# Input data
caff <- matrix(c(652, 1537, 598, 242, 36, 46, 38, 21, 218, 327, 106, 67),
              nrow = 3, dimnames = list(c("Married", "PMarried", "Single"), c("0", "1-150", "151-300", ">300")), byrow=TRUE)
addmargins(caff)
```

## 2. Determine whether Caffeine consumption and Marital status are independent
```{r}
chisq.test(caff)
chisq1 <- chisq.test(caff)
chisq1$expected
```

All expected counts are bigger than 5. We fail to reject $H_0$ since the p-value is lower than $\alpha$, `chisq1$p.value` < 0.05. Therefore, caffeine consumption and marital status are dependent, i.e. women quantity of coffe consumption is dependant on marital status.

## 3. If there is dependence between Caffeine consumption and Marital status, what is the conclusion?
```{r}
mosaicplot(caff, shade = TRUE)
```

Married women are under-represented when not consuming caffeine.
Previously married women are under-represented when consuming coffee in a low quantity (0-150), but over-represented when consuming a lot of coffee (151-300+).
Single women are over-represented when not consuming coffee (0) and under-represented when drinking considerably a lot of coffee (151-300).