---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 4D - KFM


```{r "setup", include=FALSE}
## set working directory
require("knitr")
```
The dataset kfm contains measurements of the newborn babies, their mother and milk consumption

Variable name | Description
------------- | -------------
dl.milk | amount of breast milk (dl)
sex | gender of body
weight | baby weight (kg)
ml.suppl | amount of milk supplement (ml)
mat.weight | mothers weight (kg)
mat.height | mothers height (cm)

## 1. Make apropriate plots of the birthweight in order to check wether the weight is normally distributed

```{r}
# Read data
df <- read.table("kfm.txt", header=TRUE, row.names=1)
df <- df[,-1]
head(df)

# Visual comparison of histogram with normal pdf
par(mfrow=c(1,2))
hist1 <- hist(df$weight, freq = FALSE, main="Histogram of birthweight", xlab="Birthweight (kg)")
x_range <- seq(min(df$weight), max(df$weight), by = 0.01)
lines(x_range, dnorm(x_range, mean(df$weight), sd(df$weight)), lw = 3, col = "red")
color1 <- rgb(1,0,0, alpha = 0.5)
# polygon(c(seq(3.5,4,0.05),4,3.5), c(dnorm(seq(3.5,4,0.05), mean(x), sd(x)),0,0), col = color1, border = color1)

# Visual comparison of ecdf with normal cdf
plot(ecdf(df$weight), main="ECDF vs. CDF of birthweight")
lines(x_range, pnorm(x_range, mean(df$weight), sd(df$weight)), lw = 3, col = "red")
abline(v=hist1$breaks)

# Could've added qqplot to check for normality
qqPlot()
```

From the visual inspection it seems that birthweight is normally distributed.

## 2. Make a $\chi^2$-test to test if the birthweights for the babies can be assumed normally distributed

```{r}
shapiro.test(df$weight)

# Get expected counts for all histogram bins
breaks <- hist1$breaks
prob <- pnorm(c(-Inf, breaks[-c(1, length(breaks))], Inf), mean(df$weight), sd(df$weight))
prob <- diff(prob)
expected <- length(df$weight)*prob

# Expected values are below 5, we need to adjust the bins

# Chi Square test for goodness-of-fit
counts <- hist1$counts
chi_obs <- sum((counts - expected)^2 / expected)
p_value <- 1 - pchisq(chi_obs, df = length(hist1$mids) - 3) 

# Expected values

```

Birthweight is normally distributed since p-valye is greather than $\alpha$, i.e. we fail to reject $H_0$ since `r p_value` > 0.05.