---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 3A - Kali


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
```


The dataset kali contains measurements of the content of kali (K2O) for four different productions

Variable name | Description
------------- | -------------
production | production unit
kali | content of kali


## 1. Use a non-parametric test to examine if the content of kali depends on the different productions

Start by loading and plotting the data

```{r}
df <- read.table("kali.txt", header=TRUE)
kali <- df$kali
production <- df$production
boxplot(kali ~ production, xlab = "Production Unit", ylab = "Content of kali", col = c(2,3,4,6), main="Content of kali by production unit")
plot(as.numeric(production), kali, col = c(rep(c(2,3,4,6), each = 6)), xlab="Production Unit", ylab="Content of kali", main="Content of kali by production unit", pch=19)
```

Non-parametric test

```{r}
kruskal.test(kali~production)
```

Distributions are the same since p-value is less than $\alpha$ = 0.05, i.e. we cannot reject $H_0$.

## 2. Use a one-way ANOVA to examine if the content of kali depends on the different productions

```{r}
# Calculate one-way ANOVA
lm1 <- lm(kali~production, df)
anova(lm1)

# Checking model assumptions
par(mfrow = c(1,2))
plot(lm1, col = df$production, which = 1:2, pch=19)
```

One-way ANOVA shows that there's statistical significance of the content of kali depending on the production of different productions; the means of the different groups are the same, i.e. p > $\alpha$=0.05. The graph also shows that there's similar variance of both groups and their residuals are normally distributed.

## 3. If the content of kali depends on the different productions, which of the production(s) yield the highest content?  

```{r}
summary(lm1)
```

Production b, since the intercept is the sum of production a + production b, i.e. $10.75+8.15=18.9$.