---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 3D - Fish


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
```

To study the influence of ocean grazers on regeneration rats of seaweed in the intertidal zone, a
researcher scraped rock plots free of seaweed and observed the degree of regeneration when certain
types of seaweed-grazing animals were denied access. The grazers were limpets (L), small fishes (f)
and large fishes (F). A plot was taken to be a square rock surface, 100 cm on each side. Each plot
received one of six treatments, named here by which grazers were allowed access.

symbol| Description
------------- | -------------
LfF | All three grazers were allowed access
fF | Limpets were excluded by surrounding the plot with caustic paint
Lf | Large fish were excluded by covering the plot with a course net
f | Limpets and large fish were excluded
L | Small and large fish were excluded by covering the plot with a fine net
C | Control: limpets, small fish and large fish were all excluded

Because the intertidal zone is a highly variable environment, the researcher applied the treatments
in eight blocks of twelve plots each. Within each block she randomly assigned treatments to plots so
that each treatment was applied to two plots. The blocks covered a wide range of tidal conditions

Symbol | Description
------------- | -------------
Block 1 | just below high tide level, exposed to heavy surf
Block 2 | just below high tide level, protected from the surf
Block 3 | midtide, exposed
Block 4 | midtide, protected
Block 5 | just above low tide level, exposed
Block 6 | just above low tide level, protected
Block 7 | on near-vertical rock wall, midtide level, protected
Block 8 | on near-vertical rock wall, above low tide level, protected

## **Variables** 

Variable name | Description
------------- | -------------
cover | Percentage of regenerated seaweed
block | Blocks for different tidal situations
treat | Treatment, i.e. what fishes are excluded

## 1. Make appropriate plots to investigate whether treatments or blocks have any influence on the percentage of regenerated seawed

Start by loading and inspecting data
```{r}
grazer <- read.table("fishgrazer.txt", header=TRUE)
grazer$block <- factor(grazer$block)
```

Informative plots  

```{r}
# Boxplot
par(mfrow = c(1,2))
boxplot(grazer$cover~grazer$block, grazer, xlab = "Block", ylab = "Percentage of regenerated seaweed", col = c(2,3,4,5,6,7,8,9))
boxplot(grazer$cover~grazer$treat, grazer, xlab = "Treat", ylab = "Percentage of regenerated seaweed", col = c(2,3,4,5,6,7))

# Scatter
par(mfrow = c(1,2))
plot(as.numeric(grazer$block), grazer$cover, pch=19, xlab = "Block", ylab = "Percentage of regenerated seaweed")
plot(as.numeric(grazer$treat), grazer$cover, pch=19, xlab = "Treat", ylab = "Percentage of regenerated seaweed")

# Scatter with block and treat
par(mfrow=c(1,1))
plot(as.numeric(grazer$block), grazer$cover, col = as.numeric(grazer$block)+1, pch = as.numeric(grazer$treat)+1, xlab = "Block", ylab = "Percentage of regenerated seaweed")
#legend("center", legend = c("Boys 7th", "Boys 8th", "Girls 7th", "Girls 8th"), col = c(2,3,4,5,6,7,8,9), pch = c(2,3,4,5,6,7))

# General scatter
par(mfrow=c(1,1))
plot(grazer$cover, col = as.numeric(grazer$block)+1, pch = as.numeric(grazer$treat)+1, xlab = "Block", ylab = "Percentage of regenerated seaweed")
#legend("topright", legend = c("Boys 7th", "Boys 8th", "Girls 7th", "Girls 8th"), col = c(2,3,4,5,6,7,8,9), pch = c(2,3,4,5,6,7))

# With log transformation
grazer$logcover <- log(grazer$cover)

# Boxplot with log transformation
par(mfrow = c(1,2))
boxplot(grazer$logcover~grazer$block, grazer, xlab = "Block", ylab = "Percentage of regenerated seaweed", col = c(2,3,4,5,6,7,8,9))
boxplot(grazer$logcover~grazer$treat, grazer, xlab = "Treat", ylab = "Percentage of regenerated seaweed", col = c(2,3,4,5,6,7))

# Scatter
par(mfrow = c(1,2))
plot(as.numeric(grazer$block), grazer$logcover, pch=19, xlab = "Block", ylab = "Percentage of regenerated seaweed")
plot(as.numeric(grazer$treat), grazer$logcover, pch=19, xlab = "Treat", ylab = "Percentage of regenerated seaweed")

# General scatter with log transformation
par(mfrow=c(1,1))
plot(grazer$logcover, col = as.numeric(grazer$block)+1, pch = as.numeric(grazer$treat)+1, xlab = "Block", ylab = "Percentage of regenerated seaweed")
```

From the plots, we can argue that when more grazers are present on the blocks, the seaweed regenerate less. On the contrary, the tidal activity seems to be better on the seaweed with mid-tide, while low-tide and high-tide seem harsher on the regeneration, i.e. the seaweed regenerates more with mid-tide than low- and high-tide.

## 2. Test the model using a 5% significance level
Start with a full model with interactions

```{r}
# Linear model of cover given block and treat
lm1 <- lm(grazer$cover~grazer$block*grazer$treat, grazer)

# Two-way ANOVA of block and treat
anova(lm1)

# Linear model of block and treat
summary(lm1)

# Model test of block and treat
par(mfrow=c(2,2))
plot(lm1, col = grazer$block, which = 1:2, pch=19)
plot(lm1, col = grazer$treat, which = 1:2, pch=19)

# Linear model of cover given block and treat With log of cover
lm2 <- lm(grazer$logcover~grazer$block*grazer$treat, grazer)

# Two-way ANOVA of log-transformed block and treat
anova(lm2)

# Linear model of log-transformed block and treat
summary(lm2)

# Model test of log-transformed block and treat
par(mfrow=c(2,2))
plot(lm2, col = grazer$block, which = 1:2, pch=19)
plot(lm2, col = grazer$treat, which = 1:2, pch=19)
```

## 3. Is there evidence of any interaction effects?

Only for block 7 with grazer LfF. But in general no, the effects seems to be independent.