---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 3C - Filter


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
```


The data are from a statement by Texaco, Inc. to the Air and Water Pollution Subcommittee of the
Senate Public Works Committee on June 26, 1973. Mr. John McKinley, President of Texaco, cited
the Octel filter, developed by Associated Octel Company as effective in reducing pollution. However,
questions had been raised about the effects of pollution filters on aspects of vehicle performance,
including noise levels. He referred to data presented in the datafile associated with this story as
evidence that the Octel filter was at least as good as a standard silencer in controlling vehicle
noise levels.
The dataset filter constitute a 3-way factorial experiment with 3 replications. The factors are type of
filter (2 types), vehicle size (3 sizes), and side of car (two sides).

Variable name | Description
------------- | -------------
NOISE | Noise level reading (decibels)
SIZE| Vehicle size: 1 small 2 medium 3 large
TYPE | 1 standard silencer 2 Octel filter
SIDE | 1 right side 2 left side of car


## 1. Determine whether size, type and side influence on the noise level by doing a graphical comparison

Start by loading and converting data to right class
```{r}
# Loading data and converting to factors
filter <- read.table("filter.txt", header=TRUE)
filter$SIZE <- factor(filter$SIZE)
levels(filter$SIZE) <- c("small", "medium", "large")
filter$TYPE <- factor(filter$TYPE)
levels(filter$TYPE) <- c("standard", "octel")
filter$SIDE <- factor(filter$SIDE)
levels(filter$SIDE) <- c("right", "left")

# Summary of the data
summary(filter)

# Structure of the data
str(filter)
```

Visual inspection of data
```{r}
# Boxplot
par(mfrow = c(1,3))
boxplot(filter$NOISE~filter$SIZE, filter, xlab = "Size", ylab = "Noise", col = c(2,3,4))
boxplot(filter$NOISE~filter$TYPE, filter, xlab = "Type", ylab = "Noise", col = c(7,8))
boxplot(filter$NOISE~filter$SIDE, filter, xlab = "Side", ylab = "Noise", col = c(5,6))

# Scatter plot
par(mfrow = c(1,3))
plot(as.numeric(filter$TYPE), filter$NOISE, pch=19)
plot(as.numeric(filter$SIZE), filter$NOISE, pch=19)
plot(as.numeric(filter$SIDE), filter$NOISE, pch=19)

# Scatter plot combined
par(mfrow=c(1,1))
plot(as.numeric(filter$TYPE), filter$NOISE, col = as.numeric(filter$SIDE)+1, 
     pch = as.numeric(filter$SIZE)+1, xlab = "Type", ylab = "Noise")
legend("center", legend = c("RightSmall", "LeftSmall", "RightMedium", "LeftMedium", "RightLarge", "LeftLarge"), 
       col = c(2,3), pch = c(2,2,3,3,4,4))
```
## 2. Determine whether size, type and side influence on the noise level by the appropriate statistical analysis
```{r}
# Linear Models
lmsize <- lm(filter$NOISE~filter$SIZE, filter)
lmtype <- lm(filter$NOISE~filter$TYPE, filter)
lmside <- lm(filter$NOISE~filter$SIDE, filter)
lmall <- lm(filter$NOISE~filter$SIZE+filter$TYPE+filter$SIDE, filter)

# One-way ANOVA of Size
anova(lmsize)
# Linear Model of Size
summary(lmsize)

# One-way ANOVA of Type
anova(lmtype)
# Linear Model of Type
summary(lmtype)

# One-way ANOVA of Side
anova(lmside)
# Linear Model of Side
summary(lmside)

# Three-way ANOVA of Size, Type, and Side
anova(lmall)
# Linear model of Size, Type, and Side
summary(lmall)
```

Size is statistically significant, i.e. means of the groups are the same.
Type is not statistically significant (octel), i.e. means of the groups are not the same.
Side is not statistically significant (left) , i.e. means of the groups are not the same.
For the three-way anova, side is not statistically significant, i.e. variable side doesn't have the same mean in the groups AND it doesn't affect the noise levels.

We could investigate a new model without the side variable.

```{r}
lmall_side <- lm(filter$NOISE~filter$SIZE+filter$TYPE, filter)
# Three-way ANOVA without Side
anova(lmall_side)
# Linear Model of Size and Type
summary(lmall_side)
```

Model looks better and all groups have the same mean.

## 3. Are there any interaction effects between size and type?
```{r}
# Visualize interaction
par(mfrow=c(1,1))
interaction.plot(filter$SIZE, filter$TYPE, filter$NOISE, ylab="Mean of NOISE", xlab="Size", main="Interactions between Type and Size", col=c(2,3))

# Linear Model
lmall_int <- lm(filter$NOISE~filter$SIZE*filter$TYPE, filter)

# Two-way ANOVA
anova(lmall_int)
# Linear Model of Size and Type
summary(lmall_int)
```

Size small for type octel and interaction between size large and type octel are not statistically significant. That means, for size, that it doesn't matter if the type of the filter is standard or octel but the size is the important feature. While the interaction of size large and type octel doesn't affect the model, we thus reject $H_0$.