---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 3B - Fertilizer


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
```

In a pilot study two different fertilizers were tested on three different fields. The yield (compared to
a reference) is given in the table below 

Fertilizer/Field | a  | b |  c |   
---- | ---- |  ---- | ---- |   
a | 10.3 |-4.95 | -6
b | 10.3 | -4.95 | -11

## 1. Type the data into R so that appropriate analysis can be carried out

```{r}
df <- data.frame(Fertilizer=c(rep("a",3),rep("b",3)), Field=rep(c("a","b","c"),2), Yield=c(10.3,-4.95, -6, 10.3, -4.95, -11))
fertilizer <- df$Fertilizer
field <- df$Field
yield <- df$Yield
```

## 2. Determine whether Fertilizer and/or Field influence on the yield
Start by making a one-way anova

```{r}
# Plot of the data
par(mfrow = c(1,3))
boxplot(yield~fertilizer, df, xlab = "Fertilizer", ylab = "Distance", col = c(2,3))
boxplot(yield~field, df, xlab = "Field", ylab = "Distance", col=c(4,5,6))
plot(as.numeric(fertilizer), yield, col = as.numeric(fertilizer)+1, 
     pch = as.numeric(field), xlab = "Fertilizer", ylab = "Yield")
legend("center", legend = c("a", "b", "c"), 
       col = c(2,2,2,3,3,3), pch = c(1,2,3,1,2,3))

# Calculate one-way ANOVA for fertilizer
lm1a <- lm(yield~fertilizer, df)
anova(lm1a)

# Calculate one-way ANOVA for field
lm1b <- lm(yield~field, df)
anova(lm1b)

# Checking model assumptions
par(mfrow = c(2,2))
plot(lm1a, col = fertilizer, which = 1:2, pch=19)
plot(lm1b, col = field, which = 1:2, pch=19)

# Two-way Anova
# lm2 <- lm(yield~fertilizer+field, df)
# anova(lm2)
# par(mfrow = c(1,2))
# plot(lm2, col = df$field, which = 1:2, pch=19)
```

From the One-Way ANOVA, fertilizer isn't statistically significant since the p-value is > $\alpha$ = 0.05, i.e. we reject the null hypothesis $H_0$ that the means of the groups are not the same. Field, on the other hand, is statistically significant, i.e. we fail to reject the null hypothesis $H_0$, so the means of the groups are the same.

Interpret the linear models
```{r}
# Fertilizer Linear Model
summary(lm1a)
# Field Linear Model
summary(lm1b)
```

From the linear models, we can conclude that both fertilizers aren't statistically significant, i.e. we reject the null hypothesis. However, for the field data, all three fields are statistically significant, being field a the one with the highest yield and field c the lower.

## 3. Is it possible to test for interaction effects between Fertilizer and Field?

```{r}

lm2 <- lm(yield~fertilizer*field, df)
# One-way ANOVA with interactions between fertilizer and field
anova(lm2)
# Fertilizer and Field interactions linear model
summary(lm2)
```

No, since we have one observation by factor combination. However, we can test for combinations of fields.

```{r}
# Combining fields into "ab" and "c"
df$Field2 <- df$Field
levels(df$Field2) <- rep(c(rep("ab",2), "c"),2)
field2 <- df$Field2
lm3 <- lm(yield~fertilizer*field2, df)
# One-way ANOVA of fertilizer and field interactions
anova(lm3)
# Fertilizer and Field interactions linear model
summary(lm3)
```

Interactions aren't statistically significant.