---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2D - Cheese


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
require("tree")
```


As cheddar cheese matures, a variety of chemical processes take place. The taste of matured cheese
is related to the concentration of several chemicals in the final product. In a study of cheddar cheese
from the LaTrobe Valley of Victoria, Australia, samples of cheese were analyzed for their chemical
composition and were subjected to taste tests. Overall taste scores were obtained by combining the
scores from several tasters.

Variable name | Description
------------- | -------------
case | sample number
taste | subjective taste test score, obtained by combining the scores of several tasters
acetic | Natural log of concentration of acetic acid
h2s | Natural log of concentration of hydrogen sulfide
lactic | Concentration of lactic acid


## 1. Use scatterplots, correlation, and simple regression to examine the relationships among the individual variables.
Start by loading data 

```{r}
cheese <- read.table("cheese.txt", header=TRUE)
cheese <- subset(cheese,select= taste:Lactic)
```

Visually investigate the relation between the variables:

```{r}
# Plot between variables and correlation matrix
plot(cheese, panel=panel.smooth)
cor(cheese)

# Simple linear regression models
lmacetic <- lm(cheese$taste~cheese$Acetic)
lmh2s <- lm(cheese$taste~cheese$H2S)
lmlactic <- lm(cheese$taste~cheese$Lactic)

# Summary of simple linear regression models
summary(lmacetic)
summary(lmh2s)
summary(lmlactic)
```

Intercept of H2S isn't statistically significant.

## 2. Why do you think acetic and h2s has been transformed?

Because acetic and h2s are concentrations which probably behave as a natural logarithm. By transforming the values of these variables we can "see" better how the variables relate to each other and with the dependent variable.

## 3. What happens when you run a regression model with all the independent variables in the model?

```{r}
lmall <- lm(cheese$taste~cheese$Acetic+cheese$H2S+cheese$Lactic)
summary(lmall)
par(mfrow=c(2,2))
plot(lmall)
```

It looks like acetic acid concentration isn't statistically significant in the model.

## 4. What model would you prefer for prediction?
```{r}
#Code here
```



## 5. Predict the 'taste' of a cheese where (log) acetic is 5.3, (log) h2s is 8.0 and lactic is 3.0


```{r}
# Code here
```


## 6. Could there be some problems with the above prediction?

```{r}

# Code here
```
