---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2C - Process


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
require("tree")
```

The dataset process contains measurements of air flow, water temperature, and acid concentration of
a process loss.

Variable name | Description
------------- | -------------
loss | loss from process
airflow | air flow
watertemp | temperature of water
acidconc | concentration of acid


## 1. Determine whether air flow, temperature of water, or concentration of acid influence on the process loss by a graphical comparison  
Start by loading the data
```{r}
process <- read.table("process.txt", header=TRUE)

plot(process, panel=panel.smooth)
```

## 2. Determine whether air flow, temperature of water or concentration of acid influence on the process loss by analysing each variable using simple linear regression  

Making simple regression for each variable

```{r}
lmair <- lm(process$loss~process$airflow)
lmwater <- lm(process$loss~process$watertemp)
lmacid <- lm(process$loss~process$acidconc)

summary(lmair)
summary(lmwater)
summary(lmacid)
```

Plot the scatter plots

```{r}
plot(process$loss~process$airflow, xlab="Airflow", ylab="Process Loss", main="Process loss related to airflow", col=1, pch=16)
abline(lmair)
plot(process$loss~process$watertemp, xlab="Water Temperature", ylab="Process Loss", main="Process loss related to water temperature", col=1, pch=16)
abline(lmwater)
plot(process$loss~process$acidconc, xlab="Acid Concentration", ylab="Process Loss", main="Process loss related to acid concentration", col=1, pch=16)
abline(lmacid)
```

It appears like acid concentration doesn't have an influence in process loss because of the p-value in the linear regression and the visual inspection.

## 3. Determine whether air flow, temperature of water or concentration of acid influence on the process loss using multiple linear regression  

```{r}
lmall <- lm(process$loss~process$airflow+process$watertemp+process$acidconc)
summary(lmall)
par(mfrow=c(2,2))
plot(lmall)
```

## 4. Is there evidence of multicollinearity?

```{r}
cor(process)
coplot(process$loss ~ process$watertemp | process$watertemp, process, panel = panel.smooth)
model <- tree(process$loss ~ ., process)
plot(model)
text(model)
```

It looks like airflow and water temperature are collinear/positiviely correlated.

## 5. Plot the residuals and analyse the results. Which x-variable should be removed if we want to reduce the model?


```{r}
par(mfrow=c(2,2))
plot(lmall)

# New model without acid concentration
lmall2 <- lm(process$loss~process$airflow+process$watertemp)
summary(lmall2, correlation=TRUE)

# New model without acid concentration and without the outlier
process2 <- process[-21,]
lmall3 <- lm(process2$loss~process2$airflow+process2$watertemp)
par(mfrow=c(2,2))
plot(lmall3)
summary(lmall3, correlation=TRUE)
```

We should remove acid concentration and the outlier.