---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2C - Process


```{r "setup", include=FALSE}
## set working directory
require("knitr")
```

The dataset process contains measurements of air flow, water temperature, and acid concentration of
a process loss.

Variable name | Description
------------- | -------------
loss | loss from process
airflow | air flow
watertemp | temperature of water
acidconc | concentration of acid


## 1. Determine whether air flow, temperature of water, or concentration of acid influence on the process loss by a graphical comparison  
Start by loading the data
```{r}
# Code here
```

## 2. Determine whether air flow, temperature of water or concentration of acid influence on the process loss by analysing each variable using simple linear regression  

Making simple regression for each variable

```{r}
# Code here
```


## 3. Determine whether air flow, temperature of water or concentration of acid influence on the process loss using multiple linear regression  

```{r}
# Code here
```

## 4. Is there evidence of multicollinearity?

```{r}
#Code here
```


## 5. Plot the residuals and analyse the results. Which x-variable should be removed if we want to reduce the model?


```{r}
# Code here
```

