---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2B - Brain


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
```


The dataset brainweight contains measurements of the weight of both brain and body for different
mammals

Variable name | Description
------------- | -------------
art | type of mammal
body| weigth of body (kg)
brain | weight of brain (gram)


## 1. Make a scatterplot of body against brain. Do you see any correlation?

Start by loading the data

```{r}
brain <- read.table("brainweight.txt", header=TRUE)
```

Plot body against brain
```{r}
plot(brain$body~brain$brain, xlab="Weight of brain (g)", ylab="Weight of body (kg)", main="Correlation of body against brain weight", col=1, pch=16)
```

## 2. Make a log transform of both body and brain. Make a scatterplot of the transformed variables. Compute the correlation.

```{r}
brain$logbody <- log(brain$body)
brain$logbrain <- log(brain$brain)
plot(brain$logbody~brain$logbrain, xlab="Log of weight of brain (g)", ylab="Log of weight of body (kg)", main="Correlation of body against brain weight, Log Transformed", col=1, pch=16)
```

## 3. Fit a regression model between log(body) and log(brain)

```{r}
logbody <- brain$logbody
logbrain <- brain$logbrain
lmlog <- lm(logbody~logbrain, data=brain)
```

## 4. Is there any outlier in the data?  

```{r}
summary(lmlog)
par(mfrow=c(2,2))
plot(lmlog)

brain2 <- brain[-34,]
logbody <- brain2$logbody
logbrain <- brain2$logbrain
lmlog2 <- lm(logbody~logbrain, data=brain2)
summary(lmlog2)
par(mfrow=c(2,2))
plot(lmlog2)
```

## 5. How would you evaluate the fit of the model?

```{r}
# Estimate, Standard Error, t-value, and multiple R-squared


# Plot 95% confidence intervals for the regression line
par(mfrow = c(1,1))
logbrain.new <- seq(min(logbrain), max(logbrain), by=0.1)
pred.logbrain <- data.frame(logbrain = logbrain.new)
range.pred <- predict(lmlog2, newdata = pred.logbrain, interval = "confidence")
plot(logbody~logbrain, brain, pch = 19)
lines(logbrain.new, range.pred[,1], lw = 2)
lines(logbrain.new, range.pred[,2], col = 2, lw = 2)
lines(logbrain.new, range.pred[,3], col = 2, lw = 2)

# Plot 95% prediction intervals using function matlines
range.pred <- predict(lmlog2, newdata = pred.logbrain, interval = "prediction")
matlines(logbrain.new, range.pred[,2:3], lty = 2, col = 3, lw = 2)

# Add legend
legend("topleft", legend = c("confidence interval","prediction interval"), 
       col = c(2,3), lty = c(1,2))
```

