---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2D - Cheese


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
require("tree")
```


As cheddar cheese matures, a variety of chemical processes take place. The taste of matured cheese
is related to the concentration of several chemicals in the final product. In a study of cheddar cheese
from the LaTrobe Valley of Victoria, Australia, samples of cheese were analyzed for their chemical
composition and were subjected to taste tests. Overall taste scores were obtained by combining the
scores from several tasters.

Variable name | Description
------------- | -------------
case | sample number
taste | subjective taste test score, obtained by combining the scores of several tasters
acetic | Natural log of concentration of acetic acid
h2s | Natural log of concentration of hydrogen sulfide
lactic | Concentration of lactic acid


## 1. Use scatterplots, correlation, and simple regression to examine the relationships among the individual variables.
Start by loading data 

```{r}
cheese <- read.table("cheese.txt", header=TRUE)
cheese <- subset(cheese,select= taste:Lactic)
```

Visually investigate the relation between the variables:

```{r}
# Plot between variables and correlation matrix
plot(cheese, panel=panel.smooth)
cor(cheese)

# Simple linear regression models
lmacetic <- lm(cheese$taste~cheese$Acetic)
lmh2s <- lm(cheese$taste~cheese$H2S)
lmlactic <- lm(cheese$taste~cheese$Lactic)

# Summary of simple linear regression models
summary(lmacetic)
summary(lmh2s)
summary(lmlactic)

# Scatter plot with linear regression line
plot(cheese$taste~cheese$Acetic, main="Cheese taste vs. acetic acid concentration", xlab="Acetic Acid Concentration", ylab="Taste Score",pch=19)
abline(lmacetic)
plot(cheese$taste~cheese$H2S, main="Cheese taste vs. hydrogen sulfide concentration", xlab="Hydrogen Sulfide Concentration", ylab="Taste Score",pch=19)
abline(lmh2s)
plot(cheese$taste~cheese$Lactic, main="Cheese taste vs. lactic acid concentration", xlab="Lactic Acid Concentration", ylab="Taste Score",pch=19)
abline(lmlactic)
```

Intercept of H2S isn't statistically significant.

## 2. Why do you think acetic and h2s has been transformed?

Because acetic and h2s are concentrations which probably behave as a natural logarithm. By transforming the values of these variables we can "see" better how the variables relate between each other and with the dependent variable.

## 3. What happens when you run a regression model with all the independent variables in the model?

```{r}
taste <- cheese$taste
Acetic <- cheese$Acetic
H2S <- cheese$H2S
Lactic <- cheese$Lactic
lmall <- lm(taste~Acetic+H2S+Lactic)
#lmall <- lm(cheese$taste~cheese$Acetic+cheese$H2S+cheese$Lactic)
summary(lmall, correlation=TRUE)
par(mfrow=c(2,2))
plot(lmall)
```

It looks like acetic acid concentration isn't relevant in the model. Multiple R-squared is better than individual linear regression models.

## 4. What model would you prefer for prediction?
```{r}
lmall_minusAcetic <- lm(taste~H2S+Lactic)
summary(lmall_minusAcetic)
par(mfrow=c(2,2))
plot(lmall_minusAcetic)
AIC(lmacetic, lmh2s, lmlactic, lmall, lmall_minusAcetic, k = 2)
```

I'd choose the multiple linear regression model without Acetic.

## 5. Predict the 'taste' of a cheese where (log) acetic is 5.3, (log) h2s is 8.0 and lactic is 3.0


```{r}
taste.new <- c(5.3, 8.0, 3.0)
pred.cheese <- data.frame(Acetic = taste.new[1], H2S = taste.new[2], Lactic=taste.new[3])
range.pred <- predict(lmall, newdata = pred.cheese)
range.pred <- unname(range.pred)
```


## 6. Could there be some problems with the above prediction?

```{r}
plot(cheese$taste~cheese$Acetic, ylim=c(0,70), pch=19)
points(taste.new[1],range.pred, col="red", pch=19)
plot(cheese$taste~cheese$H2S, ylim=c(0,70), pch=19)
points(taste.new[2],range.pred, col="red", pch=19)
plot(cheese$taste~cheese$Lactic, xlim=c(0,3), ylim=c(0,70), pch=19)
points(taste.new[3],range.pred, col="red", pch=19)
```
