---
output:
  pdf_document: default
  html_document: default
---

## 02441 Applied Statistics and Statistical Software 
# Exercise 2A - Sport


```{r "setup", include=FALSE}
## set working directory
require("knitr")
require("car")
require("ggplot2")
```

This dataset contains the gold medal performances in the men's long jump, high jump and discus throw
for the modern Olympic games from 1900 to 1984. Data are also provided for the 1968 Olympics, but
one may wish to omit them from the analyses because that Olympics was quite different from later
ones. $\alpha$ > chiquibeibi

Variable name | Description
------------- | -------------
highjump | winning result in high jum
discusthrow | winning result in discus throw
longjump | winning result in long jump
year | year with reference at 1900

## 1. Make appropriate plots for the different sport disciplines. Is there a tendency?
Start by loading data

```{r}
sport <- read.table("sport.txt", header=TRUE)

# First remove the results from 1896 and change the date to the reference
sport <- sport[-1,]
sport$year <- sport$year+1900
summary(sport)
```

Plotting the different sport disciplines

```{r}
# Boxplot
par(mfrow=c(1,3))
boxplot(sport$high.jump, main="High Jump", col=2)
boxplot(sport$Discus.Throw, main="Discus Throw", col=3)
boxplot(sport$long.jump, main="Long Jump", col=4)

# Line plot
par(mfrow=c(3,1))
plot(sport$year, sport$high.jump, type="l", xlab="Year", ylab="Winning results", main="High Jump", col=2)
plot(sport$year, sport$Discus.Throw, type="l", xlab="Year", ylab="Winning results", main="Discus Throw", col=3)
plot(sport$year, sport$long.jump, type="l", xlab="Year", ylab="Winning results", main="Long Jump", col=4)

# Scatter plot
par(mfrow=c(1,3))
plot(sport$year, sport$high.jump, xlab="Year", ylab="Winning results", main="High Jump", col=2, pch=16)
plot(sport$year, sport$Discus.Throw, xlab="Year", ylab="Winning results", main="Discus Throw", col=3, pch=16)
plot(sport$year, sport$long.jump, xlab="Year", ylab="Winning results", main="Long Jump", col=4, pch=16)
```

It looks like there's a positive tendency for all three sports.

## 2. It has been suggested that the Mexico City Olympics in 1968 saw unusually good track and field performances, possibly because of the high altitude. To investigate this question we must establish some expected performance level. How would you do this?  

Fit linear model with the observation for long.jump 1968. 

```{r}
lmLongJump <- lm(sport$long.jump~sport$year, data=sport)

summary(lmLongJump)

par(mfrow = c(2,2))
plot(lmLongJump)
```

Plot the linear model with observation 1968 for long jump

```{r}
# Build model again only for long jump
lj <- sport$long.jump
year <- sport$year
lmLongJump <- lm(lj~year, data=sport)

# Plot 95% confidence intervals for the regression line
par(mfrow = c(1,1))
year.new <- seq(min(year), max(year), by=0.1)
#year.new <- c(2004)
pred.sport <- data.frame(year = year.new)
range.pred <- predict(lmLongJump, newdata = pred.sport, interval = "confidence")
plot(lj ~ year, sport, pch = 19)
lines(year.new, range.pred[,1], lw = 2)
lines(year.new, range.pred[,2], col = 2, lw = 2)
lines(year.new, range.pred[,3], col = 2, lw = 2)

# Plot 95% prediction intervals using function matlines
range.pred <- predict(lmLongJump, newdata = pred.sport, interval = "prediction")
matlines(year.new, range.pred[,2:3], lty = 2, col = 3, lw = 2)

# Add legend
legend("topleft", legend = c("confidence interval","Prediction interval"), 
       col = c(2,3), lty = c(1,2))
```

Fit linear model without observation for long.jump 1968. 

```{r}
# Remove observation
sport2 <- sport[-15,]

# Linear Model
lj <- sport2$long.jump
year <- sport2$year
lmLongJump <- lm(lj~year, data=sport2)

summary(lmLongJump)

par(mfrow = c(2,2))

plot(lmLongJump)
```

Plot the linear model without observation 1968

```{r}
# Plot 95% confidence intervals for the regression line
par(mfrow = c(1,1))
year.new <- seq(min(year), max(year), by=0.1)
pred.sport <- data.frame(year = year.new)
range.pred <- predict(lmLongJump, newdata = pred.sport, interval = "confidence")
plot(lj ~ year, sport2, pch = 19)
lines(year.new, range.pred[,1], lw = 2)
lines(year.new, range.pred[,2], col = 2, lw = 2)
lines(year.new, range.pred[,3], col = 2, lw = 2)

# Plot 95% prediction intervals using function matlines
range.pred <- predict(lmLongJump, newdata = pred.sport, interval = "prediction")
matlines(year.new, range.pred[,2:3], lty = 2, col = 3, lw = 2)

# Add legend
legend("topleft", legend = c("confidence interval","Prediction interval"), 
       col = c(2,3), lty = c(1,2))
```

## 3. Could you predict the winning result in long jump for the Olympic games in Greece 2004? 
```{r}
year.new <- c(2004)
pred.sport <- data.frame(year = year.new)
range.pred <- unname(predict(lmLongJump, newdata = pred.sport))
```

The predicted value is 342.62 for the long jump sport in Greece 2004.